---
title: "prediction_XER_pctIMP_OE"
author: "C. Emi Fergus"
date: "2023-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## % Impervious surface in watershed riparian buffer

There are only nine sites in the Xeric ecoregion with % impervious surface in stream buffer > 5%. Most NRSA sites have less than 1% impervious surface in the riparian buffer. 

But for those nine sites, we want to see how reducing % impervious surface would improve stream condition.

For sites with impervious surface in riparian buffer, will create scenarios where impervious surface is reduced.

```{r,echo=FALSE, warning=FALSE, message=FALSE}
remove(list=ls())

library(tidyverse)
library(lavaan)
library(dplyr)
library(ggplot2)
library(ggpmisc)
library(ggpubr)
library(broom)
library(GGally)

library(kableExtra)
library(knitr)
library(stringr)

library(sf) # Simple Features works well with Tidyverse
library(mapview)
library(RColorBrewer)

devtools::load_all()
library(SEMscenarios)
```


```{r,echo=FALSE, warning=FALSE}
##############
## READ PROCESSED NRSA DATA USED IN MODEL
# PROCESSED - all 0809 and only new sites from later surveys
#  VISITS 1 and 2 n = 4578 w/ 321 vars
dat_org<-read.csv("C:/Users/efergus/SEMscenarios/data_processed/Compiled/nrsa081318_nonresampled_VISIT_12.csv")


# PROCESSED DATA VISIT_NO=1 ONLY n = 4389
dat_proc<-dat_org%>%
  filter(VISIT_NO==1)

###############
## PROCESS DATA DROPPING MISSING PROTOCOL
dat_proc$PROTOCOL<-as.factor(dat_proc$PROTOCOL)

# n = 4371
dat_proc<- dat_proc%>%
  drop_na(PROTOCOL)%>%
  filter(PROTOCOL=="BOATABLE"|PROTOCOL=="WADEABLE")

# DROP NOPHAB class from REALM
dat_proc$PROTOCOL<-droplevels(dat_proc$PROTOCOL)

# TRANSFORM IMPERVIOUS SURFACE (NLCD
dat_proc<-dat_proc%>%
  mutate(asin_PCTIMP_WS = asin(sqrt(PCTIMP_WS/100)),
         asin_PCTIMP_WsRp100 = asin(sqrt(PCTIMP_WsRp100/100)),
         asin_PCTIMP_CAT = asin(sqrt(PCTIMP_CAT/100)),
         asin_PCTIMP_CATRP100 = asin(sqrt(PCTIMP_CATRP100/100)))

#########################
## SUBSET BY MANUAL ECOREGION
# XER n = 342
xer<-dat_proc%>%
  filter(AG_ECO9=="XER")%>%
  drop_na(LOE_QLow_cl)

# SCALE CUMULATIVE PRECIPITATION in XER
xer$PSUMPY_SY_WS_sc<-scale(xer$PSUMPY_SY_WS)

# WADEABLE n = 200
xer_w <- xer %>%
  filter(PROTOCOL=="WADEABLE")


#############
## READ SEM OUTPUT PROCESSED W/COEFFICIENT LABELS UNSTANDARDIZED NON-BAYESIAN version 3
coef_proc<-read.csv("C:/Users/efergus/SEMscenarios/Routput/Scenario_modeling/SEM_output/XERw_m3_OE_coef_unstd.csv")
#head(coef_proc)

# Grab label and coefficient estimate to call in equations
test3<-coef_proc[,3:4]
# Widen table - make coefficient labels columns and estimates the row
coef_use<-pivot_wider(test3,names_from=coeff_name,values_from = est)#

```


## Examine Predicted vs Observed relationship for sites with impervious surface
Based on coefficients from path analysis model, estimated what O/E values would be using the dataset the model was trained on. We can compare predicted and observed values to see how well the model fit.  
```{r, echo=FALSE , warning=FALSE}
##########################################
###### EXAMINING OE PATH MODEL USING DATASET MODEL WAS TRAINED ON
# Subset data % impervious >5% n = 9 sites
xer_impv<-xer_w%>%
  filter(PCTIMP_WsRp100>5)

oe.sem.predict <- coef_use$b9.1 + coef_use$b9.2*xer_impv$LRBS_use + coef_use$b9.3*xer_impv$L_NTL +
  coef_use$b9.4*xer_impv$L_SULF + coef_use$b9.5*xer_impv$LOE_Qbkf_cl + coef_use$b9.6*xer_impv$L_NABD_NrmStorWs_ratio +
  coef_use$b9.7*xer_impv$L_STRM_POWER

#  ADD PREDICTED VALUES TO XER subset
xer_train_pred <- xer_impv%>%
  mutate(OE_SCORE_pred=oe.sem.predict)

xer_train_pred<-xer_train_pred%>%
  mutate(resid_oe=OE_SCORE-OE_SCORE_pred)

# RMSE - Calculate RMSE (sqrt(mean (obs-pred)^2))
rmse_train_oe<-round(sqrt(mean(xer_train_pred$resid_oe)^2),3)
rmse_oe_lab<-paste("RMSE =",rmse_train_oe)

# COMPARE OBSERVED VS MLR PREDICTED (n=111)
OE_corr_train<-ggplot(xer_train_pred,aes(x=OE_SCORE, y=OE_SCORE_pred))+ #PCTFOR_WsRp100
  #stat_poly_line()+
  #stat_poly_eq(use_label(c("eq","R2")))+
  geom_point()+
  geom_abline(intercept=0,slope=1)+
  geom_text(x=0.7,y=0.97, label=rmse_oe_lab)+
  xlim(0,1.5)+
  ylim(0,1.5)+
  xlab("Observed OE")+
  ylab("Predicted OE")+
  ggtitle("Predicted OE vs Observed with 1:1 line")
OE_corr_train

```

For the nine sites with >5% impervious surface, predicted O/E is greater than observed. Suggests that there are other things impairing these sites that aren't well captured by the proximal drivers but may be associated with impervious surface (i.e., other pollutants).

The RMSE is 0.16, indicating that average distance between the predicted O/E and observed O/E is 0.16 units. This difference seems like it would be problematic because a 16% loss in taxa richness might be important to the ecosystem.

## SCENARIO % IMPERVIOUS SURFACE
Reduce impervious surface for sites with values >5%
Scenario 1: Reduce by 0%
Scenario 2: Reduce by 5% (X% - 5%)
Scenario 3: Reduce impervious surface by 50%

Bug O/E ~ 1.18 + 0.06(LRBS_use) + -0.14(L_NTL) + -0.08(L_SULF) + 0.05(LOE_Qbkf_cl) + 0.04(dam) + 0.12(Specific strm pwr)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# CREATE HYPOTHETICAL % Impervious
# Scenario where decrease impervious surface by 5 units
xer_mod<-xer_impv%>%
  mutate(asin_PCTIMP.s2  = asin_PCTIMP_WsRp100 - 0.225513)%>%
  mutate(asin_PCTIMP.s2.raw  = (sin(asin_PCTIMP.s2 ))^2) %>%
  mutate(asin_PCTIMP.s3 = asin_PCTIMP_WsRp100-(asin_PCTIMP_WsRp100*0.5))%>%
  mutate(asin_PCTIMP.s3.raw = (sin(asin_PCTIMP.s3 ))^2)

summary(xer_mod$asin_PCTIMP.s2.raw)
summary(xer_mod$asin_PCTIMP.s3.raw)

# SCENARIO 1 - OBSERVED Impervious
asin_PCTIMP.predict.1 <- xer_mod$asin_PCTIMP_WsRp100
asin_PCTIMP.predict.1.raw<-xer_mod$PCTIMP_WsRp100

#SCENARIO 2 - Decrease Impervious by 5 units
asin_PCTIMP.predict.2 <- xer_mod$asin_PCTIMP.s2
asin_PCTIMP.predict.2.raw<-xer_mod$asin_PCTIMP.s2.raw

# SCENARIO 3 - Decrease Impervious by 50%
asin_PCTIMP.predict.3 <- xer_mod$asin_PCTIMP.s3
asin_PCTIMP.predict.3.raw<-xer_mod$asin_PCTIMP.s3.raw

# PREDICTED DISTRIBUTIONS FOR EXOGENOUS DRIVERS
# VALUES IN DATASET
#imp.pred.1 <- xer_mod$asin_PCTIMP_WsRp100
dam.pred.1 <- xer_mod$L_NABD_NrmStorWs_ratio
hag.pred.1 <- xer_mod$W1_HAG
nohag.pred.1 <- xer_mod$W1_HNOAG
rnat.pred.1 <- xer_mod$asin_PCTFORGRS_CATRP100
precip.pred.1 <-xer_mod$PSUMPY_SY_WS_sc
phdi.pred.1<-xer_mod$drought_mean

# ENDOGENOUS DRIVERS BUT NOT ALTERED BY Impervious
strmpwr.pred.1 <- xer_mod$L_STRM_POWER
xcmgw.pred.1 <- xer_mod$Lpt01_XCMGW
loebflow.pred.1 <- xer_mod$LOE_Qbkf_cl

#############
## DRIVERS OF O/E ALTERED by IMPERVIOUS
# SCENARIO 1 - no change in impervious surface
loelflow.predict.1 <-coef_use$b3.1 + coef_use$b3.2*asin_PCTIMP.predict.1 +
  coef_use$b3.3*precip.pred.1 + coef_use$b3.4*dam.pred.1 + 
  coef_use$b3.5*rnat.pred.1 + coef_use$b3.6*hag.pred.1 +
  coef_use$b3.7*nohag.pred.1

evap.predict.1 <- coef_use$b5.1 + coef_use$b5.2*loelflow.predict.1 +
  coef_use$b5.3*loebflow.pred.1 +
  coef_use$b5.4*precip.pred.1 + coef_use$b5.5*dam.pred.1 +
  coef_use$b5.6*strmpwr.pred.1

rbs.predict.1<- coef_use$b6.1 + coef_use$b6.2*asin_PCTIMP.predict.1 +
  coef_use$b6.3*loelflow.predict.1 + coef_use$b6.4*loebflow.pred.1 +
  coef_use$b6.5*hag.pred.1 + coef_use$b6.6*rnat.pred.1 +
  coef_use$b6.7*precip.pred.1 + coef_use$b6.8*strmpwr.pred.1
rbs.predict.1.raw<-10^rbs.predict.1

tn.predict.1<- coef_use$b7.1 + coef_use$b7.2*asin_PCTIMP.predict.1 +
  coef_use$b7.3*xcmgw.pred.1 + coef_use$b7.4*loelflow.predict.1 +
  coef_use$b7.5*precip.pred.1
tn.predict.1.raw<-10^tn.predict.1 -0.01
  
sulf.predict.1<- coef_use$b8.1 + coef_use$b8.2*asin_PCTIMP.predict.1+
  coef_use$b8.3*evap.predict.1 + coef_use$b8.4*nohag.pred.1 +
  coef_use$b8.5*rnat.pred.1 + coef_use$b8.6*loelflow.predict.1 +
  coef_use$b8.7*precip.pred.1 + coef_use$b8.8*phdi.pred.1
sulf.predict.1.raw<-10^sulf.predict.1


oe.predict.1 <- coef_use$b9.1 + coef_use$b9.2*rbs.predict.1 +
  coef_use$b9.3*tn.predict.1 + coef_use$b9.4*sulf.predict.1 +
   coef_use$b9.5*loebflow.pred.1  + coef_use$b9.6*dam.pred.1 +
  coef_use$b9.7*strmpwr.pred.1

############
# SCENARIO #2
loelflow.predict.2 <-coef_use$b3.1 + coef_use$b3.2*asin_PCTIMP.predict.2 +
  coef_use$b3.3*precip.pred.1 + coef_use$b3.4*dam.pred.1 + 
  coef_use$b3.5*rnat.pred.1 + coef_use$b3.6*hag.pred.1 +
  coef_use$b3.7*nohag.pred.1

evap.predict.2 <- coef_use$b5.1 + coef_use$b5.2*loelflow.predict.2 +
  coef_use$b5.3*loebflow.pred.1 +
  coef_use$b5.4*precip.pred.1 + coef_use$b5.5*dam.pred.1 +
  coef_use$b5.6*strmpwr.pred.1

rbs.predict.2<- coef_use$b6.1 + coef_use$b6.2*asin_PCTIMP.predict.2 +
  coef_use$b6.3*loelflow.predict.2 + coef_use$b6.4*loebflow.pred.1 +
  coef_use$b6.5*hag.pred.1 + coef_use$b6.6*rnat.pred.1 +
  coef_use$b6.7*precip.pred.1 + coef_use$b6.8*strmpwr.pred.1
rbs.predict.2.raw<-10^rbs.predict.2

tn.predict.2<- coef_use$b7.1 + coef_use$b7.2*asin_PCTIMP.predict.2 +
  coef_use$b7.3*xcmgw.pred.1 + coef_use$b7.4*loelflow.predict.2 +
  coef_use$b7.5*precip.pred.1
tn.predict.2.raw<-10^tn.predict.2 -0.01
  
sulf.predict.2<- coef_use$b8.1 + coef_use$b8.2*asin_PCTIMP.predict.2+
  coef_use$b8.3*evap.predict.2 + coef_use$b8.4*nohag.pred.1 +
  coef_use$b8.5*rnat.pred.1 + coef_use$b8.6*loelflow.predict.2 +
  coef_use$b8.7*precip.pred.1 + coef_use$b8.8*phdi.pred.1
sulf.predict.2.raw<-10^sulf.predict.2

oe.predict.2 <- coef_use$b9.1 + coef_use$b9.2*rbs.predict.2 +
  coef_use$b9.3*tn.predict.2 + coef_use$b9.4*sulf.predict.2 +
  coef_use$b9.5*loebflow.pred.1  + coef_use$b9.6*dam.pred.1 +
  coef_use$b9.7*strmpwr.pred.1

############
# SCENARIO #3
loelflow.predict.3 <-coef_use$b3.1 + coef_use$b3.2*asin_PCTIMP.predict.3 +
  coef_use$b3.3*precip.pred.1 + coef_use$b3.4*dam.pred.1 + 
  coef_use$b3.5*rnat.pred.1 + coef_use$b3.6*hag.pred.1 +
  coef_use$b3.7*nohag.pred.1

evap.predict.3 <- coef_use$b5.1 + coef_use$b5.2*loelflow.predict.3 +
  coef_use$b5.3*loebflow.pred.1 +
  coef_use$b5.4*precip.pred.1 + coef_use$b5.5*dam.pred.1 +
  coef_use$b5.6*strmpwr.pred.1

rbs.predict.3<- coef_use$b6.1 + coef_use$b6.2*asin_PCTIMP.predict.3 +
  coef_use$b6.3*loelflow.predict.3 + coef_use$b6.4*loebflow.pred.1 +
  coef_use$b6.5*hag.pred.1 + coef_use$b6.6*rnat.pred.1 +
  coef_use$b6.7*precip.pred.1 + coef_use$b6.8*strmpwr.pred.1
rbs.predict.3.raw<-10^rbs.predict.3

tn.predict.3<- coef_use$b7.1 + coef_use$b7.2*asin_PCTIMP.predict.3 +
  coef_use$b7.3*xcmgw.pred.1 + coef_use$b7.4*loelflow.predict.3 +
  coef_use$b7.5*precip.pred.1
tn.predict.3.raw<-10^tn.predict.3 -0.01
  
sulf.predict.3<- coef_use$b8.1 + coef_use$b8.2*asin_PCTIMP.predict.3+
  coef_use$b8.3*evap.predict.3 + coef_use$b8.4*nohag.pred.1 +
  coef_use$b8.5*rnat.pred.1 + coef_use$b8.6*loelflow.predict.3 +
  coef_use$b8.7*precip.pred.1 + coef_use$b8.8*phdi.pred.1
sulf.predict.3.raw<-10^sulf.predict.3

oe.predict.3 <- coef_use$b9.1 + coef_use$b9.2*rbs.predict.3 +
  coef_use$b9.3*tn.predict.3 + coef_use$b9.4*sulf.predict.3 +
  coef_use$b9.5*loebflow.pred.1  + coef_use$b9.6*dam.pred.1 +
  coef_use$b9.7*strmpwr.pred.1
```


Density distributions of predicted values under % Impervious surface scenarios
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#######################
## CREATE DATAFRAMES OF PREDICTED VALUES
scenario.1<-data.frame(asin_PCTIMP.predict.1,asin_PCTIMP.predict.1.raw,
                       loelflow.predict.1,
                       rbs.predict.1,rbs.predict.1.raw, 
                       tn.predict.1,tn.predict.1.raw,
                       sulf.predict.1,sulf.predict.1.raw,
                       oe.predict.1) #xfcnat.predict.1.raw,

scenario.2<-data.frame(asin_PCTIMP.predict.2,asin_PCTIMP.predict.2.raw,
                       loelflow.predict.2,
                       rbs.predict.2,rbs.predict.2.raw, 
                       tn.predict.2,tn.predict.2.raw,
                       sulf.predict.2,sulf.predict.2.raw,
                       oe.predict.2)

scenario.3<-data.frame(asin_PCTIMP.predict.3,asin_PCTIMP.predict.3.raw,
                       loelflow.predict.3,
                       rbs.predict.3,rbs.predict.3.raw, 
                       tn.predict.3,tn.predict.3.raw,
                       sulf.predict.3,sulf.predict.3.raw,
                       oe.predict.3)

# ADD COLUMNS TO INDIVIDUAL SCENARIO DATASETS TO INDICATE WHICH SCENARIO THEY BELONG TO
scenario.1<-scenario.1%>%
  mutate(scenario=1)
scenario.2<-scenario.2%>%
  mutate(scenario=2)
scenario.3<-scenario.3%>%
  mutate(scenario=3)

# change column names
cols_names<-c("asin_PCTIMP.predict","PCTIMP.predict","LOElflow.predict",
              "Lrbs.predict","Rbs.predict","Ltn.predict","tn.predict",
              "Lsulf.predict","sulf.predict","oe.predict","scenario")
names(scenario.1)<- cols_names
names(scenario.2)<- cols_names
names(scenario.3)<- cols_names

# SCENARIOS GRP 1 - Decrease by 5 units
scenario_all<-rbind(scenario.1,scenario.2,scenario.3) #,scenario.4
scenario_all$scenario<-as.factor(scenario_all$scenario)



####################
## DENSTIY PLOTS OF PREDICTED VALUES
library(ggplot2)
#mycols <- c("#7570b3","#1b9e77","#d95f02","")
mycols<- c("#d7191c","#fdae61","#2c7bb6") #light blue "#abd9e9",

# Get group mean SCENARIO
mu_oe <- scenario_all%>%
  #filter(!is.na(rbs.predict.1.raw))%>%
  group_by(scenario)%>%
  summarize(mean_oe = mean(oe.predict))
mu_tn <- scenario_all%>%
  #filter(!is.na(rfor.pred.1.raw))%>%
  group_by(scenario)%>%
  summarize(mean_tn = mean(tn.predict))

##################
# PREDICTED O/E DISTRIBUTIONS
dens_OE<-ggplot(scenario_all, aes(x=oe.predict, fill=scenario, colour=scenario))+
  geom_density(alpha=0.3)+
  scale_fill_manual(values=mycols)+
  scale_color_manual(values=mycols)+
  geom_vline(data=mu_oe, aes(xintercept=mean_oe, colour=scenario),
             linetype="dashed")+
  theme_bw()+
  theme(plot.title = element_text(face="plain",size=11, hjust=0.5))+
  ggtitle("Predicted O/E")+
  xlab("Predicted O/E")+
  theme(legend.position = "bottom")

dens_OE

# PREDICTED TN DISTRIBUTIONS - NEED TO BACKTRANSFORM
dens_TN<-ggplot(scenario_all, aes(x=tn.predict, fill=scenario, colour=scenario))+
  geom_density(alpha=0.3)+
  scale_fill_manual(values=mycols)+
  scale_color_manual(values=mycols)+
  geom_vline(data=mu_tn, aes(xintercept=mean_tn, colour=scenario),
             linetype="dashed")+
  theme_bw()+
  theme(plot.title = element_text(face="plain",size=11, hjust=0.5))+
  ggtitle("Predicted TN")+
  xlab("Predicted TN (mg/L)")+
  theme(legend.position = "bottom")

dens_TN

```

## SUBSET BY SITES WHERE IMPERVIOUS SURFACE CHANGED (Sites w/%IMPV > 5%)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########################
## COMBINE PREDICTED WITH OBSERVED
################
## Combine scenario predictions with NRSA dataset reduced to sites 
xer_mod_red<- xer_mod%>%
  select(SITE_ID,VISIT_NO,YEAR,UNIQUE_ID,LAT_DD83,LON_DD83,RT_MASTER,
         OE_SCORE,LOE_QLow_cl,LRBS_use,L_NTL,L_SULF,L_STRM_POWER,asin_PCTIMP_WsRp100,WGT_TP)
xer_pred1<-cbind(xer_mod_red,scenario.1)
xer_pred2<-cbind(xer_mod_red,scenario.2)
xer_pred3<-cbind(xer_mod_red,scenario.3)

# Calculate difference in predicted O/E and observed
xer_pred1<- xer_pred1%>%
  mutate(OE_SCORE_diff=oe.predict-OE_SCORE)
xer_pred2<- xer_pred2%>%
  mutate(OE_SCORE_diff=oe.predict-OE_SCORE)
xer_pred3<- xer_pred3%>%
  mutate(OE_SCORE_diff=oe.predict-OE_SCORE)

xer_prediction<-rbind(xer_pred1,xer_pred2,xer_pred3)
xer_prediction$scenario<- as.factor(xer_prediction$scenario)

```


## Calculate mean O/E by scenario

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########################
## Estimate mean values
################
df_impv<-xer_prediction%>%
  group_by(scenario)%>%
  summarise(mean_OE = mean(oe.predict))

df_impv_median<-xer_prediction%>%
  group_by(scenario)%>%
  summarise(median_OE = median(oe.predict))

kable(df_impv, caption = "Mean Predicted Bug O/E",digits=3)%>%
  kable_styling()
  
```


Boxplots of predicted O/E distributions

```{r, echo=FALSE,message=FALSE,warning=FALSE}
# Annotate text in boxplot indicating number of sites by TN class
dat_text <- data.frame(
  label=c(paste("\u0394","O/E = 0.108"),paste("\u0394","O/E = 0.087")),
  x=c(2,3),
  y=c(0.9),
  scenario="1"
)

pred_OE_impv <- ggplot(xer_prediction, aes(x=scenario, y=oe.predict,fill=scenario))+ #weight = WGT_TP
  geom_boxplot()+
  scale_fill_manual(values=mycols)+
  geom_jitter(alpha=0.1)+
  geom_text(data=dat_text,
            mapping=aes(x=x,y=y,label=label))+
  #geom_text(x="1",y=1.25, aes(label=label),data=ann_text)+
  theme(axis.text.x = element_text(angle=65, hjust=1,size=10),
        axis.title.x = element_blank())+
  xlab("Scenario")+
  ylab("Predicted O/E")
pred_OE_impv

```

## Population estimates

See Dumelle et al. 2023 for tips using spsurvey package
```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(spsurvey)

# SCENARIO 1 n=135 - sites with TN greater than GOod/Fair condition
xer_prediction_s1<-xer_prediction%>%
  filter(scenario=="1")%>%
  filter(!is.na(WGT_TP))

# Convert dataframe to a simple features (sf) object
# https://stackoverflow.com/questions/29736577/how-to-convert-data-frame-to-spatial-coordinates/45484314#45484314
xer_pred_s1_sf <- st_as_sf(x=xer_prediction_s1,
                           coords=c("LON_DD83","LAT_DD83"),
                           crs=4269) # Albers equal area EPSG = 5070

cont_est_s1 <-cont_analysis(
  xer_pred_s1_sf,
  siteID = "SITE_ID",
  vars="oe.predict",
  weight="WGT_TP"
)

# Mean estimates of Scenario 1 predicted O/E
cont_est_s1$Mean
# Estimate of oe.predict for scenario 1 = 0.54

# CDF estimates
#plot(cont_est_s1$CDF)

# Scenario 2
xer_prediction_s2<-xer_prediction%>%
  filter(scenario=="2")%>%
  filter(!is.na(WGT_TP))

xer_pred_s2_sf <- st_as_sf(x=xer_prediction_s2,
                           coords=c("LON_DD83","LAT_DD83"),
                           crs=4269) # Albers equal area EPSG = 5070

cont_est_s2 <-cont_analysis(
  xer_pred_s2_sf,
  siteID = "SITE_ID",
  vars="oe.predict",
  weight="WGT_TP"
)

# Mean estimates of Scenario 2 predicted O/E
cont_est_s2$Mean
# Estimate of oe.predict for scenario 2 = 0.646


## Scenario 3
xer_prediction_s3<-xer_prediction%>%
  filter(scenario=="3")%>%
  filter(!is.na(WGT_TP))

xer_pred_s3_sf <- st_as_sf(x=xer_prediction_s3,
                           coords=c("LON_DD83","LAT_DD83"),
                           crs=4269) # Albers equal area EPSG = 5070

cont_est_s3 <-cont_analysis(
  xer_pred_s3_sf,
  siteID = "SITE_ID",
  vars="oe.predict",
  weight="WGT_TP"
)

# Mean estimates of Scenario 3 predicted O/E
cont_est_s3$Mean


#################
## COMBINE CDF dataframes
s1_cdf<-cont_est_s1$CDF%>%
  mutate(scenario="scenario_1")

s2_cdf<-cont_est_s2$CDF%>%
  mutate(scenario="scenario_2")

s3_cdf<-cont_est_s3$CDF%>%
  mutate(scenario="scenario_3")

cdf_all_highTN <- rbind(s1_cdf,s2_cdf,s3_cdf)

# CDF for WMT & XER
oe_cdf<-ggplot(cdf_all_highTN, aes(x=Value))+
  stat_ecdf(aes(color=scenario),
            geom="step", size=1.5)+
  theme(legend.position="bottom")+
  xlab("O/E predicted")+
  ggtitle("XER population O/E \nscenario reduce % impervious surface by 5 units")

oe_cdf

# PRINT CDF
#tiff(filename="C:/Users/efergus/SEMNRSA/inst/analysis_prediction/Routput/F#igures/Population_OE_scenario_v5_2023_1103.tiff",
#     width=5, height=4, units="in", res=200)
#oe_cdf
#dev.off()
```


Estimate population of Xeric sites in % impervious surface classes

```{r,echo=FALSE,message=FALSE,warning=FALSE}

# Create class of impervious surface
xer_w_red<-xer_w%>%
  mutate(impv_class = case_when(
    asin_PCTIMP_WsRp100 >0.225513 ~ ">5% impervious",
    asin_PCTIMP_WsRp100 <=0.225513 ~ "<5% impervious"))%>%
  filter(!is.na(WGT_TP))
table(xer_w_red$impv_class)

xer_w_sf <- st_as_sf(x=xer_w_red,
                           coords=c("LON_DD83","LAT_DD83"),
                           crs=4269) # Albers equal area EPSG = 5070

#################
## ESTIMATE POPULATION OF XER STREAMS BY IMpervious CLASS

cat_ests <- cat_analysis(
  xer_w_sf,
  siteID = "SITE_ID",
  vars = "impv_class",
  weight = "WGT_TP"
)
cat_ests
# Is this % stream km in the XER by impervious surface class?
cat_ests_red<-cat_ests%>%
  filter(Category %in% c(">5% impervious","<5% impervious"))

cat_ests_red$Category<-ordered(cat_ests_red$Category, levels=c("<5% impervious",">5% impervious"))

col_palette <-c("#91bfdb","#ffffbf") #"#fc8d59"
#Set font
windowsFonts(AR=windowsFont("Arial"))

XER_impv_class<-ggplot(cat_ests_red, aes(x=Category, y=Estimate.P,fill=Category))+
  geom_bar(stat="identity",position = position_dodge())+
  geom_errorbar(aes(ymax=Estimate.P + MarginofError.P,ymin=Estimate.P -MarginofError.P ),
                width=0.2)+
  scale_fill_manual(values=col_palette)+
  geom_text(aes(label=paste0(round(Estimate.P,0),"%")),vjust=-0.2)+
  #theme_bw(base_size=12)+
  theme(plot.title = element_text(family = "AR",face="plain",size=14, hjust=0.5),
        axis.text.x = element_text(family = "AR", angle=45, hjust=1,size=12),#, colour=c(rep("#8c510a",3),rep("#bf812d",1),rep("#dfc27d",3),rep("#c7eae5",2),rep("#80cdc1",1),rep("#c7eae5",1),rep("#35978f",2),rep("#01665e",1))),
        axis.text.y = element_text(family = "AR", size=12),
        axis.title.y = element_text(family="AR"), #element_blank(),#
        strip.text.x = element_text(family="AR", size=12),
        panel.grid.major =  element_line(colour = NA),
        panel.grid.minor=element_line(colour = NA))+
  xlab("")+
  ylab("% of Stream length")+
  ggtitle("Xeric impervious surface class") +
  coord_flip()
XER_impv_class


```

