---
title: "prediction_XER_OE_LNTL_v4"
author: "C. Emi Fergus"
date: "2023-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scenario predictions using path analysis coefficients
 Alan H suggested taking XERIC sites with poor TN (25%; TN > 529ug/L)
  What would happen to these streams if you reduced TN at each site
  by 10%, 20%, and/or 30%.  I think those are more realistic in terms of
  what might actually be accomplished by some kind of nutrient controls or
  BMPs in a watershed.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lavaan)
library(dplyr)
library(ggplot2)
library(ggpmisc)
library(ggpubr)
library(broom)

library(grid)
library(gridExtra)

library(kableExtra)
library(knitr)
library(stringr)

devtools::load_all()
library(SEMscenarios)
```

```{r,echo=FALSE, message=FALSE, warning=FALSE}
# LOAD DATA
##############
## READ PROCESSED NRSA DATA USED IN MODEL
# PROCESSED - all 0809 and only new sites from later surveys
#  VISITS 1 and 2 n = 4578 w/ 321 vars
dat_org<-read.csv("C:/Users/efergus/SEMscenarios/data_processed/Compiled/nrsa081318_nonresampled_VISIT_12.csv")

# PROCESSED DATA VISIT_NO=1 ONLY n = 4389
dat_proc<-dat_org%>%
  filter(VISIT_NO==1)

###############
## PROCESS DATA DROPPING MISSING PROTOCOL
dat_proc$PROTOCOL<-as.factor(dat_proc$PROTOCOL)

# n = 4371
dat_proc<- dat_proc%>%
  drop_na(PROTOCOL)%>%
  filter(PROTOCOL=="BOATABLE"|PROTOCOL=="WADEABLE")

# DROP NOPHAB class from REALM
dat_proc$PROTOCOL<-droplevels(dat_proc$PROTOCOL)

#########################
## SUBSET BY MANUAL ECOREGION
# XER n = 342
xer<-dat_proc%>%
  filter(AG_ECO9=="XER")%>%
  drop_na(LOE_QLow_cl)

# SCALE CUMULATIVE PRECIPITATION in XER
xer$PSUMPY_SY_WS_sc<-scale(xer$PSUMPY_SY_WS)

# WADEABLE n = 200
xer_w <- xer %>%
  filter(PROTOCOL=="WADEABLE")

# POOR TN SITES (TN > 529 ug/L or > -0.28) based on NRSA 13-14 report thresholds
xer_mod<-xer_w%>%
mutate(TN_class = case_when(
  L_NTL > -0.28 ~ "TN_Poor",
  L_NTL <= -0.28 ~ "other"))
table(xer_mod$TN_class)

#############
## READ SEM OUTPUT PROCESSED W/COEFFICIENT LABELS UNSTANDARDIZED NON-BAYESIAN
coef_proc<-read.csv("C:/Users/efergus/SEMscenarios/Routput/Scenario_modeling/SEM_output/XERw_m3_OE_coef_unstd.csv")

# Grab label and coefficient estimate to call in equations
coef_labels<-coef_proc[,3:4]
# Widen table - make coefficient labels columns and estimates the row
coef_use<-pivot_wider(coef_labels,names_from=coeff_name,values_from = est)#

```

## CREATE HYPOTHETICAL TN VALUES FOR NRSA SITES greater than reference
```{r, echo=FALSE, message=FALSE, echo=FALSE}
# SUBSET OF XER Sites with Poor TN
# Decrease untransformed TN by 10%, 20%, and 30%
xer_tn<-xer_mod%>%
  filter(TN_class =="TN_Poor")%>%
  mutate(NTL_s1 = NTL_RESULT-(NTL_RESULT*0.1))%>%
  mutate(LNTL_s1 = log10(NTL_s1))%>%
  mutate(NTL_s2=NTL_RESULT-(NTL_RESULT*0.2))%>%
  mutate(LNTL_s2 = log10(NTL_s2))%>%
  mutate(NTL_s3=NTL_RESULT-(NTL_RESULT*0.3))%>%
  mutate(LNTL_s3 = log10(NTL_s3))

head(xer_tn[,c(31,324,326,328)],10)

```

# Set scenario values and predict bug O/E

Bug O/E ~ 1.18 + 0.06(LRBS_use) + -0.14(L_NTL) + -0.08(L_SULF) + 0.05(LOE_Qbkf_cl) + 0.04(dam) + 0.12(Specific strm pwr)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# SCENARIO 0 - OBSERVED TN
L_NTL.predict.0 <- xer_tn$L_NTL
tn.predict.0.raw<-xer_tn$NTL_RESULT

#SCENARIO 1 - TN decrease by 10%
L_NTL.predict.1 <- xer_tn$LNTL_s1
tn.predict.1.raw<-xer_tn$NTL_s1

#SCENARIO 2 - TN decrease by 20%
L_NTL.predict.2 <- xer_tn$LNTL_s2
tn.predict.2.raw<-xer_tn$NTL_s2

#SCENARIO 3 - TN decrease by 30%
L_NTL.predict.3 <- xer_tn$LNTL_s3
tn.predict.3.raw<-xer_tn$NTL_s3

# PREDICTED DISTRIBUTIONS FOR EXOGENOUS DRIVERS
# VALUES IN DATASET
urb.pred.1 <- xer_tn$asin_PCTURB_CAT
dam.pred.1 <- xer_tn$L_NABD_NrmStorWs_ratio
hag.pred.1 <- xer_tn$W1_HAG
nohag.pred.1 <- xer_tn$W1_HNOAG
rnat.pred.1 <- xer_tn$asin_PCTFORGRS_CATRP100
precip.pred.1 <-xer_tn$PSUMPY_SY_WS_sc
phdi.pred.1<-xer_tn$drought_mean

# ENDOGENOUS DRIVERS BUT NOT ALTERED BY LFLOW
strmpwr.pred.1 <- xer_tn$L_STRM_POWER
xcmgw.pred.1 <- xer_tn$Lpt01_XCMGW
loebflow.pred.1 <- xer_tn$LOE_Qbkf_cl
loelflow.pred.1 <- xer_tn$LOE_QLow_cl
rbs.pred.1 <- xer_tn$LRBS_use
sulf.pred.1 <- xer_tn$L_SULF


############
# PREDICTED DISTRIBUTIONS FOR BENTHIC O/E
# SCENARIO #0
oe.predict.0 <- coef_use$b9.1 + coef_use$b9.2*rbs.pred.1 +
  coef_use$b9.3*L_NTL.predict.0 + coef_use$b9.4*sulf.pred.1 +
  coef_use$b9.5*loebflow.pred.1  + coef_use$b9.6*dam.pred.1 +
  coef_use$b9.7*strmpwr.pred.1

# SCENARIO #1
oe.predict.1 <- coef_use$b9.1 + coef_use$b9.2*rbs.pred.1 +
  coef_use$b9.3*L_NTL.predict.1 + coef_use$b9.4*sulf.pred.1 +
  coef_use$b9.5*loebflow.pred.1  + coef_use$b9.6*dam.pred.1 +
  coef_use$b9.7*strmpwr.pred.1

# SCENARIO #2
oe.predict.2 <- coef_use$b9.1 + coef_use$b9.2*rbs.pred.1 +
  coef_use$b9.3*L_NTL.predict.2 + coef_use$b9.4*sulf.pred.1 +
  coef_use$b9.5*loebflow.pred.1  + coef_use$b9.6*dam.pred.1 +
  coef_use$b9.7*strmpwr.pred.1

# SCENARIO #3
oe.predict.3 <- coef_use$b9.1 + coef_use$b9.2*rbs.pred.1 +
  coef_use$b9.3*L_NTL.predict.3 + coef_use$b9.4*sulf.pred.1 +
  coef_use$b9.5*loebflow.pred.1  + coef_use$b9.6*dam.pred.1 +
  coef_use$b9.7*strmpwr.pred.1
```

Create dataframes of predicted values
```{r, echo=FALSE, warning=FALSE}
#######################
## CREATE DATAFRAMES OF PREDICTED VALUES
scenario.0<-data.frame(L_NTL.predict.0,tn.predict.0.raw,
                       oe.predict.0)
scenario.1<-data.frame(L_NTL.predict.1,tn.predict.1.raw,
                       oe.predict.1) #xfcnat.predict.1.raw,
scenario.2<-data.frame(L_NTL.predict.2,tn.predict.2.raw,
                       oe.predict.2)
scenario.3<-data.frame(L_NTL.predict.3,tn.predict.3.raw,
                       oe.predict.3)

# ADD COLUMNS TO INDIVIDUAL SCENARIO DATASETS TO INDICATE WHICH SCENARIO THEY BELONG TO
scenario.0<-scenario.0%>%
  mutate(scenario=0)
scenario.1<-scenario.1%>%
  mutate(scenario=1)
scenario.2<-scenario.2%>%
  mutate(scenario=2)
scenario.3<-scenario.3%>%
  mutate(scenario=3)

# change column names
cols_names<-c("LNTL.predict",
              "tn.predict.raw","oe.predict","scenario")
names(scenario.0)<- cols_names
names(scenario.1)<- cols_names
names(scenario.2)<- cols_names
names(scenario.3)<- cols_names

# ROW BIND SCENARIOS ESTIMATES n = 348 obs
scenario_all<-rbind(scenario.0,scenario.1,scenario.2,scenario.3) #,scenario.4
scenario_all$scenario<-as.factor(scenario_all$scenario)

```

# Mean and Median predicted values
```{r, echo=FALSE,warning=FALSE,message=FALSE}
# Get group mean SCENARIO v1
median_oe <- scenario_all%>%
  #filter(!is.na(rbs.predict.1.raw))%>%
  group_by(scenario)%>%
  summarize(median_oe = median(oe.predict))
median_tn <- scenario_all%>%
  #filter(!is.na(rfor.pred.1.raw))%>%
  group_by(scenario)%>%
  summarize(median_tn = median(tn.predict.raw))

# Mean predictoed
mean_oe <- scenario_all%>%
  group_by(scenario)%>%
  summarize(mean_oe = mean(oe.predict))
mean_tn <- scenario_all%>%
  group_by(scenario)%>%
  summarize(mean_tn = mean(tn.predict.raw))

#Calculate difference in predicted O/E and baseline
wide<-scenario_all%>%
  select(scenario,oe.predict)%>%
  pivot_wider(names_from = scenario, values_from = oe.predict)

#df_diff<-wide%>%
#  rename(SITE_ID=1)%>%
#  rename(TN_class=2)%>%
#  rename(scenario_1 = 3)%>%
#  rename(scenario_2 = 4)%>%
#  mutate(OE_pred12_diff = scenario_2 - scenario_1)

```

# Boxplots of predicted values
```{r, echo=FALSE,warning=FALSE}
####################
## PREDICTED O/E BOXPLOTS
## DENSTIY PLOTS OF PREDICTED VALUES
mycols<- c("#a6611a","#dfc27d","#80cdc1","#018571")
# # Delta symbol
# https://stackoverflow.com/questions/27690729/greek-letters-symbols-and-line-breaks-inside-a-ggplot-legend-label
dat_text <- data.frame(
  label=c(paste("\u0394","O/E = 0.05"),paste("\u0394","O/E = 0")),
  TN_class=factor(c("hi TN","other")),
  x=c(0,1,2,3),
  y=c(1.0,1.0,1,1),
  scenario="1"
)
OE_pred_box <- ggplot(scenario_all, aes(x=scenario, y=oe.predict,fill=scenario))+
  geom_boxplot()+
  scale_fill_manual(values=mycols)+
  geom_jitter(alpha=0.2)+
  #facet_grid(cols=vars(AG_ECO9))+
  theme(axis.text.x = element_text(angle=65, hjust=1,size=10),
        axis.title.x = element_blank())+
  xlab("Scenario")+
  ylab("Predicted O/E")
OE_pred_box

#################
## TABLE OF VALUES
kable(median_oe, caption = "Median Predicted Bug O/E",digits=3)%>%
  kable_styling()

kable(mean_oe, caption = "Mean Predicted Bug O/E",digits=3)%>%
  kable_styling()

```
