---
title: "Candidate model (v3) comparisons for XER OE,MMI,EPT"
author: "C. Emi Fergus"
date: "2024-01-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(tidyverse)
library(sf)
library(spmodel)
library(lavaan)
library(semPlot)

library(ranger)
library(ggpmisc) # for lm equation
library(ggpubr)
library(dotwhisker)
library(knitr)
devtools::load_all()
library(SEMscenarios)
```

## Candidate model evaluation for macroinvertebrate O/E, MMI, and EPT 
We are interested in comparing sets of candidate models for each of the macroinvertebrate response variables (OE, MMI, EPT) separately. Candidate models included multiple linear regression, spatial linear regression, structural equation model, random forest, and random forest with spatial residuals. We evaluate candidate model predictive performance using RMSE (lower is better) and psuedo R2 values using 10-fold cross validation, and examined model coefficients (for only mlr, spatial model, & SEM).

```{r, echo=FALSE,warning=FALSE,message=FALSE}
#########################
# LOAD DATA
# Processed NRSA data - all 0809 and only new sites from later surveys
#  VISITS 1 and 2 n = 4578 w/ 321 vars
dat_org<-read.csv("C:/Users/efergus/SEMscenarios/data_processed/Compiled/nrsa081318_nonresampled_VISIT_12.csv")

# PROCESSED DATA VISIT_NO=1 WADEABLE n = 2736
dat_proc<-dat_org%>%
  filter(VISIT_NO==1)%>%
  mutate(PROTOCOL=as.factor(PROTOCOL))%>%
  drop_na(PROTOCOL)%>%
  filter(PROTOCOL=="WADEABLE")

# TRANSFORM IMPERVIOUS SURFACE; DROP MISSING QLOW n = 2584
dat_proc<-dat_proc%>%
  mutate(asin_PCTIMP_WS = asin(sqrt(PCTIMP_WS/100)),
         asin_PCTIMP_WsRp100 = asin(sqrt(PCTIMP_WsRp100/100)),
         asin_PCTIMP_CAT = asin(sqrt(PCTIMP_CAT/100)),
         asin_PCTIMP_CATRP100 = asin(sqrt(PCTIMP_CATRP100/100)))%>%
  mutate(L_OE_SCORE = log10(OE_SCORE+0.01))%>%
  drop_na(LOE_QLow_cl)%>%
  drop_na(OE_SCORE)%>%
  mutate(PSUMPY_SY_WS_sc=scale(PSUMPY_SY_WS))


##########
# XERIC n = 271 observations per unique site in NRSA surveys (08/09 + 13/14 + 18/19)
xer_org<-dat_proc%>%
  filter(AG_ECO9=="XER")

```

```{r, echo=FALSE,warning=FALSE,message=FALSE}
#########################
## CANDIDATE MODEL COMPARISONS BY MACROINVERTEBRATE RESPONSE
# LOAD Candidate model comparison parameters: RMSE and R2

# RMSE
rmse_oe<-read_csv("C:/Users/efergus/SEMscenarios/Routput/Scenario_modeling/Candidate_models/XER_v3_RMSE_OE.csv")

rmse_oe<-rmse_oe%>%
  filter(!Models=="sp_gamma")%>%
  filter(!Models=="sp_invgaus")%>%
  rename(OE=RMSE)

rmse_mmi<-read_csv("C:/Users/efergus/SEMscenarios/Routput/Scenario_modeling/Candidate_models/XER_v3_RMSE_MMI.csv")
rmse_ept<-read_csv("C:/Users/efergus/SEMscenarios/Routput/Scenario_modeling/Candidate_models/XER_v3_RMSE_EPT.csv")

rmse_mmi<-rmse_mmi%>%
  rename(MMI=RMSE)

rmse_ept<-rmse_ept%>%
  rename(EPT=RMSE)

rmse_part<-left_join(rmse_oe,rmse_mmi,join_by(Models))
rmse_all<-left_join(rmse_part,rmse_ept,join_by(Models))

knitr::kable(rmse_all, digits=3,caption="RMSE of candidate models")

############
# R2
r2_oe<-read_csv("C:/Users/efergus/SEMscenarios/Routput/Scenario_modeling/Candidate_models/XER_v3_R2_OE.csv")

r2_oe<-r2_oe%>%
  filter(!Models=="sp_gamma")%>%
  filter(!Models=="sp_invgaus")%>%
  rename(OE=R2)

r2_mmi<-read_csv("C:/Users/efergus/SEMscenarios/Routput/Scenario_modeling/Candidate_models/XER_v3_R2_MMI.csv")
r2_ept<-read_csv("C:/Users/efergus/SEMscenarios/Routput/Scenario_modeling/Candidate_models/XER_v3_R2_EPT.csv")

r2_mmi<-r2_mmi%>%
  rename(MMI=R2)

r2_ept<-r2_ept%>%
  rename(EPT=R2)

r2_part<-left_join(r2_oe,r2_mmi,join_by(Models))
r2_all<-left_join(r2_part,r2_ept,join_by(Models))

knitr::kable(r2_all, digits=2, caption="R2 of candidate models")

```

## Compare model prediction uncertainties
```{r,echo=FALSE, warning=FALSE,message=FALSE}
# https://stackoverflow.com/questions/38109501/how-does-predict-lm-compute-confidence-interval-and-prediction-interval
# Compare predictive estimates and standard error for MLR, spatial, and SEM
# Multiple linear regression (non-spatial)
 # mlr.oe <-lm(OE_SCORE~., data=train_red)# use all variables in subset except log OE
 # summary(mlr.oe)


  # Predict O/E using MLR model
 # pred.mlr<-predict(mlr.oe, test_red)

## STRUCTURAL EQUATION MODEL PREDICTION
  ## USING lavPredictY
  # PATH ANLAYSIS MODEL v3
mymodel_v3_OE <- '
Lpt01_XCMGW ~ 1 + ha3*W1_HAG + na2*W1_HNOAG + d2*L_NABD_NrmStorWs_ratio + sp3*L_STRM_POWER + p3*PSUMPY_SY_WS_sc
L_STRM_POWER ~ 1 + d3*L_NABD_NrmStorWs_ratio + ha6*W1_HAG + rn4*asin_PCTFORGRS_CATRP100
LOE_QLow_cl~ 1 + i5*asin_PCTIMP_WsRp100 + p8*PSUMPY_SY_WS_sc + d5*L_NABD_NrmStorWs_ratio + rn5*asin_PCTFORGRS_CATRP100 + ha5*W1_HAG + na4*W1_HNOAG
LOE_Qbkf_cl ~ 1 + p7*PSUMPY_SY_WS_sc + d4*L_NABD_NrmStorWs_ratio + ha4*W1_HAG + ph3*drought_mean
evap_index_sc ~ 1 + l5*LOE_QLow_cl + b3*LOE_Qbkf_cl + p6*PSUMPY_SY_WS_sc + d6*L_NABD_NrmStorWs_ratio + sp4*L_STRM_POWER
LRBS_use ~ 1 + i2*asin_PCTIMP_WsRp100 + l2*LOE_QLow_cl + b2*LOE_Qbkf_cl + ha2*W1_HAG + rn2*asin_PCTFORGRS_CATRP100 + p2*PSUMPY_SY_WS_sc + sp2*L_STRM_POWER
L_NTL ~ 1 + i3*asin_PCTIMP_WsRp100 + x2*Lpt01_XCMGW + l3*LOE_QLow_cl + p4*PSUMPY_SY_WS_sc
L_SULF ~  1 + i4*asin_PCTIMP_WsRp100 + e2*evap_index_sc + na3*W1_HNOAG + rn3*asin_PCTFORGRS_CATRP100 + l4*LOE_QLow_cl + p5*PSUMPY_SY_WS_sc + ph2*drought_mean

OE_SCORE ~ 1 + r1*LRBS_use + n1*L_NTL + su1*L_SULF + boe1*LOE_Qbkf_cl + d1*L_NABD_NrmStorWs_ratio + sp1*L_STRM_POWER

# Covariance
LOE_QLow_cl~~LOE_Qbkf_cl
L_STRM_POWER~~LOE_QLow_cl
L_SULF~~L_NTL

'

  # ROBUST ESTIMATION MAX LIKELIHOOD METHOD
fit<- sem(mymodel_v3_OE, data=xer_org,
            estimator="MLM")

#lavaan output https://lavaan.ugent.be/tutorial/inspect.html
#fitMeasures(fit)
  
# PREDICT Using lavPredictY - specify variables to include
xvars<-c("PSUMPY_SY_WS_sc","drought_mean","LRBS_use","L_NTL","L_SULF","LOE_Qbkf_cl","LOE_QLow_cl","W1_HAG","W1_HNOAG","asin_PCTIMP_WsRp100","L_NABD_NrmStorWs_ratio","L_STRM_POWER","evap_index_sc","asin_PCTFORGRS_CATRP100","Lpt01_XCMGW")

  ###########
  # SEM: PREDICT Y-VALUES
pred.sem<-lavPredictY(fit,ynames="OE_SCORE",xnames=xvars,newdata = xer_org)

semPaths(fit, title=FALSE, intercepts = FALSE, residuals =FALSE)
```

```{r echo=FALSE,message=FALSE,warning=FALSE}

xnames<-xvars
ynames<-c("OE_SCORE","L_NTL","L_SULF","LRBS_use")

# Lets look at components that are used in the predicty function
# Returns the model-implied fitted covariance matrix
Sxx<-fitted(fit)$cov[xnames, xnames]
Sxy = fitted(fit)$cov[xnames, ynames]
# Returns the mean values
mx = fitted(fit)$mean[xnames]
my = fitted(fit)$mean[ynames]

# Create matrix of x variables
Xtest = as.matrix(xer_org[,xnames])
# Center x variables by their mean
Xtest = scale(Xtest, center=mx, scale=FALSE) # center x values by subtracting from mean?

# Predicted y values based on mean of y + centered x variables times the variance covariance matrix *  
yhat = matrix(my, nrow=nrow(Xtest),ncol=length(ynames), byrow=TRUE) +
  Xtest %*% solve(Sxx) %*% Sxy

# Make predictions using the de Rooij function
# https://github.com/mjderooij/SEMpredict/blob/main/predicty.lavaan.R
pred.org <-predicty(fit,newdata=xer_org,xnames=xnames,ynames=ynames)
head(pred.org)
pred.org<-as.data.frame(pred.org)%>%
  remove_rownames() %>% #https://tibble.tidyverse.org/reference/rownames.html
  rename(rooij=OE_SCORE)
pred.OE<-as.data.frame(cbind(pred.sem,pred.org))%>%
  rename(lavaan=OE_SCORE)

# Check if values match
pred.OE <- pred.OE%>%
  mutate(check=ifelse(round(lavaan,3) == round(rooij,3), 'yes','no'))

not_match<-pred.OE%>%
  filter(check=='no')
# The values match using the different SEM prediction functions - which may be similar under the hood

```


## MLR without interaction terms 
```{r,echo=FALSE,warning=FALSE,message=FALSE}
myvars<- c("PSUMPY_SY_WS_sc", "drought_mean",
           "asin_PCTIMP_WsRp100","L_NABD_NrmStorWs_ratio",
           "L_STRM_POWER","LOE_QLow_cl","LOE_Qbkf_cl","evap_index_sc",
           "W1_HAG","W1_HNOAG","asin_PCTFORGRS_CATRP100",
           "Lpt01_XCMGW","LRBS_use","L_NTL","L_SULF","OE_SCORE")
xer_red<-xer_org%>%
  select(all_of(myvars))
summary(lm(OE_SCORE~., data= xer_red))


```


## MLR with interaction terms
Included interactions between impervious surface and TN, SO4, and summer flow
```{r,echo=FALSE,warning=FALSE,message=FALSE}
summary(lm(OE_SCORE~PSUMPY_SY_WS_sc+drought_mean+
                  asin_PCTIMP_WsRp100+L_NABD_NrmStorWs_ratio+
                  L_STRM_POWER+LOE_QLow_cl+LOE_Qbkf_cl+evap_index_sc+
                  W1_HAG+W1_HNOAG+asin_PCTFORGRS_CATRP100+
                  Lpt01_XCMGW+LRBS_use+L_NTL+L_SULF+asin_PCTIMP_WsRp100*LOE_QLow_cl+
                    asin_PCTIMP_WsRp100*L_NTL+asin_PCTIMP_WsRp100*L_SULF, data= xer_org))

```


## Dotplots of standardized coefficients
```{r, echo=FALSE,warning=FALSE,message=FALSE}
#########################
# Dot whisker plot of regression coefficients https://cran.r-project.org/web/packages/dotwhisker/vignettes/dotwhisker-vignette.html

## MACROINVERTEBRATE O/E
# Multiple linear regression (non-spatial)
# REDUCE VARIABLES TO INCLUDE IN MODEL

m1 <- xer_red %>%
  do(broom::tidy(lm(OE_SCORE~., data= xer_red))) %>% # Will "tidy" lm coefficient output into a tibble
  #rename(model=am)%>%
  relabel_predictors(c(asin_PCTIMP_WsRp100="Impervious Rp",L_NABD_NrmStorWs_ratio="Dam",
                       PSUMPY_SY_WS_sc="Precipitation",Tmax8110Ws="Max Temp",drought_mean="Drought index",
                       LOE_Qbkf_cl="OE Bankfull flow",LOE_QLow_cl="OE Summer flow",evap_index_sc="Evaporation indicator",
                       L_STRM_POWER="Specific strm power",W1_HAG="Agr index Rp", W1_HNOAG="Non-agr index Rp",
                       asin_PCTFORGRS_CATRP100="ForestGrass Rp",Lpt01_XCMGW="Riparian cover",
                       LRBS_use="Bed stability",L_NTL="TN",L_SULF="Sulfate"))%>%
  mutate(model="Multiple reg")

#m1

#################
# SPATIAL MODEL
# Make a simple features object
xer_sf <- st_as_sf(x=xer_org,
                   coords=c("LON_DD83","LAT_DD83"),
                   crs=4269)%>%
  st_transform(crs=5070)# Albers equal area EPSG = 5070 - need to project to a coordinate system where x and y units are on same scale

##########
# Spatial GLM Gaussian distribution and exponential covariance structure
m2<- xer_sf%>%
  do(broom::tidy(splm(OE_SCORE~ PSUMPY_SY_WS_sc+drought_mean+
                  asin_PCTIMP_WsRp100+L_NABD_NrmStorWs_ratio+
                  L_STRM_POWER+LOE_QLow_cl+LOE_Qbkf_cl+evap_index_sc+
                  W1_HAG+W1_HNOAG+asin_PCTFORGRS_CATRP100+
                  Lpt01_XCMGW+LRBS_use+L_NTL+L_SULF,
                xer_sf, spcov_type="exponential")))%>%
  relabel_predictors(c(asin_PCTIMP_WsRp100="Impervious Rp",L_NABD_NrmStorWs_ratio="Dam",
                       PSUMPY_SY_WS_sc="Precipitation",Tmax8110Ws="Max Temp",drought_mean="Drought index",
                       LOE_Qbkf_cl="OE Bankfull flow",LOE_QLow_cl="OE Summer flow",evap_index_sc="Evaporation indicator",
                       L_STRM_POWER="Specific strm power",W1_HAG="Agr index Rp", W1_HNOAG="Non-agr index Rp",
                       asin_PCTFORGRS_CATRP100="ForestGrass Rp",Lpt01_XCMGW="Riparian cover",
                       LRBS_use="Bed stability",L_NTL="TN",L_SULF="Sulfate"))%>%
  mutate(model="Spatial")
#m2

#########################
## STRUCTURAL EQUATION MODEL
# READ IN SEM bootstrap OE Model
sem_oe<-read.csv("C:/Users/EFergus/OneDrive - Environmental Protection Agency (EPA)/a_NLA_OE_project/Project_repository/Routput/Scenario_modeling/SEM_output/XERw_m3_OE_CI.csv")
# Modify SEM output to get dotplot
m3<- sem_oe%>%
  filter(str_detect(label,"_tot"))%>%
  rename(term=label,
         estimate=est.std,
         std.error=se,
         conf.low= ci.lower,
         conf.high=ci.upper,
         statistic=z,
         p.value=pvalue)%>%
  mutate(term = recode_factor(term, imp_tot="Impervious Rp",dam_tot="Dam",
                              precip_tot="Precipitation",phdi_tot="Drought index",
                              oebflow_tot="OE Bankfull flow",oelflow_tot="OE Summer flow",dexcess_tot="Evaporation indicator",
                              strpwr_tot="Specific strm power",
                              hag_tot="Agr index Rp", nhag_tot="Non-agr index Rp",rnat_tot="ForestGrass Rp",
                              xcmgw_tot="Riparian cover",rbs_tot="Bed stability", tn_tot="TN", sulf_tot="Sulfate"))%>%

  mutate(model="SEM")%>%
  select(term,estimate,std.error,statistic,p.value,model)

#m3

###############
## MLR with interactions
##
m4 <- xer_org %>%
  do(broom::tidy(lm(OE_SCORE~PSUMPY_SY_WS_sc+drought_mean+
                  asin_PCTIMP_WsRp100+L_NABD_NrmStorWs_ratio+
                  L_STRM_POWER+LOE_QLow_cl+LOE_Qbkf_cl+evap_index_sc+
                  W1_HAG+W1_HNOAG+asin_PCTFORGRS_CATRP100+
                  Lpt01_XCMGW+LRBS_use+L_NTL+L_SULF+asin_PCTIMP_WsRp100*LOE_QLow_cl+
                    asin_PCTIMP_WsRp100*L_NTL+asin_PCTIMP_WsRp100*L_SULF, data= xer_org))) %>% # Will "tidy" lm coefficient output into a tibble
  #rename(model=am)%>%
  relabel_predictors(c(asin_PCTIMP_WsRp100="Impervious Rp",L_NABD_NrmStorWs_ratio="Dam",
                       PSUMPY_SY_WS_sc="Precipitation",Tmax8110Ws="Max Temp",drought_mean="Drought index",
                       LOE_Qbkf_cl="OE Bankfull flow",LOE_QLow_cl="OE Summer flow",evap_index_sc="Evaporation indicator",
                       L_STRM_POWER="Specific strm power",W1_HAG="Agr index Rp", W1_HNOAG="Non-agr index Rp",
                       asin_PCTFORGRS_CATRP100="ForestGrass Rp",Lpt01_XCMGW="Riparian cover",
                       LRBS_use="Bed stability",L_NTL="TN",L_SULF="Sulfate"))%>%
  mutate(model="MLR w/interaction")
#m4


#######################
# Bring OE model output together
m_set<-rbind(m1, m2,m3,m4)
coeff_dotplot<-dwplot(m_set)+
  theme_bw()+
  ggtitle("Std. model coeff: XER wade O/E")

coeff_dotplot
#dwplot(m_set,style="distribution") # Looks crazy because very high peaks for SEM

# PRINT Coefficient Dot and whisker plot
# OE WEST
tiff(filename="C:/Users/EFergus/OneDrive - Environmental Protection Agency (EPA)/a_NLA_OE_project/Project_repository/Routput/Scenario_modeling/Figures/Dot_whisker_XER_OE.tiff",
     width=5, height = 6, units="in", res=200)
coeff_dotplot
dev.off()

```


```{r, echo=FALSE,warning=FALSE,message=FALSE}
###############################
### MMI
#########################
# Multiple linear regression (non-spatial)
m1_mmi <- xer_red %>%
  do(broom::tidy(lm(MMI_BENT_sc~ PSUMPY_SY_WS_sc+drought_mean+
                      asin_PCTIMP_WsRp100+L_NABD_NrmStorWs_ratio+
                      L_STRM_POWER+LOE_QLow_cl+LOE_Qbkf_cl+evap_index_sc+
                      W1_HAG+W1_HNOAG+asin_PCTFORGRS_CATRP100+
                      Lpt01_XCMGW+LRBS_use+L_NTL+L_SULF, data= xer_org))) %>% # Will "tidy" lm coefficient output into a tibble
  #rename(model=am)%>%
  relabel_predictors(c(asin_PCTIMP_WsRp100="Impervious Rp",L_NABD_NrmStorWs_ratio="Dam",
                       PSUMPY_SY_WS_sc="Precipitation",Tmax8110Ws="Max Temp",drought_mean="Drought index",
                       LOE_Qbkf_cl="OE Bankfull flow",LOE_QLow_cl="OE Summer flow",evap_index_sc="Evaporation indicator",
                       L_STRM_POWER="Specific strm power",W1_HAG="Agr index Rp", W1_HNOAG="Non-agr index Rp",
                       asin_PCTFORGRS_CATRP100="ForestGrass Rp",Lpt01_XCMGW="Riparian cover",
                       LRBS_use="Bed stability",L_NTL="TN",L_SULF="Sulfate"))%>%
  mutate(model="Multiple reg")

#m1_mmi

#################
# SPATIAL MODEL

##########
# Spatial GLM Gaussian distribution and exponential covariance structure
m2_mmi<- xer_sf%>%
  do(broom::tidy(splm(MMI_BENT_sc~ PSUMPY_SY_WS_sc+drought_mean+
                        asin_PCTIMP_WsRp100+L_NABD_NrmStorWs_ratio+
                        L_STRM_POWER+LOE_QLow_cl+LOE_Qbkf_cl+evap_index_sc+
                        W1_HAG+W1_HNOAG+asin_PCTFORGRS_CATRP100+
                        Lpt01_XCMGW+LRBS_use+L_NTL+L_SULF,
                      xer_sf, spcov_type="exponential")))%>%
  relabel_predictors(c(asin_PCTIMP_WsRp100="Impervious Rp",L_NABD_NrmStorWs_ratio="Dam",
                       PSUMPY_SY_WS_sc="Precipitation",Tmax8110Ws="Max Temp",drought_mean="Drought index",
                       LOE_Qbkf_cl="OE Bankfull flow",LOE_QLow_cl="OE Summer flow",evap_index_sc="Evaporation indicator",
                       L_STRM_POWER="Specific strm power",W1_HAG="Agr index Rp", W1_HNOAG="Non-agr index Rp",
                       asin_PCTFORGRS_CATRP100="ForestGrass Rp",Lpt01_XCMGW="Riparian cover",
                       LRBS_use="Bed stability",L_NTL="TN",L_SULF="Sulfate"))%>%
  mutate(model="Spatial")
#m2_mmi

#########################
## STRUCTURAL EQUATION MODEL
# READ IN SEM bootstrap Model
sem_mmi<-read.csv("C:/Users/EFergus/OneDrive - Environmental Protection Agency (EPA)/a_NLA_OE_project/Project_repository/Routput/Scenario_modeling/SEM_output/XERw_m3_MMI_CI.csv")
# Modify SEM output to get dotplot
m3_mmi<- sem_mmi%>%
  filter(str_detect(label,"_tot"))%>%
  rename(term=label,
         estimate=est.std,
         std.error=se,
         conf.low= ci.lower,
         conf.high=ci.upper,
         statistic=z,
         p.value=pvalue)%>%
  mutate(term = recode_factor(term, imp_tot="Impervious Rp",dam_tot="Dam",
                              precip_tot="Precipitation",phdi_tot="Drought index",
                              oebflow_tot="OE Bankfull flow",oelflow_tot="OE Summer flow",dexcess_tot="Evaporation indicator",
                              strpwr_tot="Specific strm power",
                              hag_tot="Agr index Rp", nhag_tot="Non-agr index Rp",rnat_tot="ForestGrass Rp",
                              xcmgw_tot="Riparian cover",rbs_tot="Bed stability", tn_tot="TN", sulf_tot="Sulfate"))%>%

  mutate(model="SEM")%>%
  select(term,estimate,std.error,statistic,p.value,model)

#m3_mmi

#######################
# Bring model output together
m_set_mmi<-rbind(m1_mmi, m2_mmi,m3_mmi)
coeff_dotplot_mmi<-dwplot(m_set_mmi)+
  theme_bw()+
  ggtitle("Std. model coeff: XER wade MMI")

coeff_dotplot_mmi
#dwplot(m_set,style="distribution") # Looks crazy because very high peaks for SEM

# PRINT Coefficient Dot and whisker plot
# MMI XER
tiff(filename="C:/Users/EFergus/OneDrive - Environmental Protection Agency (EPA)/a_NLA_OE_project/Project_repository/Routput/Scenario_modeling/Figures/Dot_whisker_XER_MMI.tiff",
     width=5, height = 6, units="in", res=200)
coeff_dotplot_mmi
dev.off()
```

```{r, echo=FALSE,warning=FALSE,message=FALSE}
###############
#### EPT
# Multiple linear regression (non-spatial)
m1_ept <- xer_red %>%
  do(broom::tidy(lm(EPT_RICH_sc~ PSUMPY_SY_WS_sc+drought_mean+
                      asin_PCTIMP_WsRp100+L_NABD_NrmStorWs_ratio+
                      L_STRM_POWER+LOE_QLow_cl+LOE_Qbkf_cl+evap_index_sc+
                      W1_HAG+W1_HNOAG+asin_PCTFORGRS_CATRP100+
                      Lpt01_XCMGW+LRBS_use+L_NTL+L_SULF, data= xer_org))) %>% # Will "tidy" lm coefficient output into a tibble
  #rename(model=am)%>%
  relabel_predictors(c(asin_PCTIMP_WsRp100="Impervious Rp",L_NABD_NrmStorWs_ratio="Dam",
                       PSUMPY_SY_WS_sc="Precipitation",Tmax8110Ws="Max Temp",drought_mean="Drought index",
                       LOE_Qbkf_cl="OE Bankfull flow",LOE_QLow_cl="OE Summer flow",evap_index_sc="Evaporation indicator",
                       L_STRM_POWER="Specific strm power",W1_HAG="Agr index Rp", W1_HNOAG="Non-agr index Rp",
                       asin_PCTFORGRS_CATRP100="ForestGrass Rp",Lpt01_XCMGW="Riparian cover",
                       LRBS_use="Bed stability",L_NTL="TN",L_SULF="Sulfate"))%>%
  mutate(model="Multiple reg")

#m1_ept

#################
# SPATIAL MODEL

##########
# Spatial GLM Gaussian distribution and exponential covariance structure
m2_ept<- xer_sf%>%
  do(broom::tidy(splm(EPT_RICH_sc~ PSUMPY_SY_WS_sc+drought_mean+
                        asin_PCTIMP_WsRp100+L_NABD_NrmStorWs_ratio+
                        L_STRM_POWER+LOE_QLow_cl+LOE_Qbkf_cl+evap_index_sc+
                        W1_HAG+W1_HNOAG+asin_PCTFORGRS_CATRP100+
                        Lpt01_XCMGW+LRBS_use+L_NTL+L_SULF,
                      xer_sf, spcov_type="exponential")))%>%
  relabel_predictors(c(asin_PCTIMP_WsRp100="Impervious Rp",L_NABD_NrmStorWs_ratio="Dam",
                       PSUMPY_SY_WS_sc="Precipitation",Tmax8110Ws="Max Temp",drought_mean="Drought index",
                       LOE_Qbkf_cl="OE Bankfull flow",LOE_QLow_cl="OE Summer flow",evap_index_sc="Evaporation indicator",
                       L_STRM_POWER="Specific strm power",W1_HAG="Agr index Rp", W1_HNOAG="Non-agr index Rp",
                       asin_PCTFORGRS_CATRP100="ForestGrass Rp",Lpt01_XCMGW="Riparian cover",
                       LRBS_use="Bed stability",L_NTL="TN",L_SULF="Sulfate"))%>%
  mutate(model="Spatial")
#m2_ept

#########################
## STRUCTURAL EQUATION MODEL
# READ IN SEM bootstrap EPT Model
sem_ept<-read.csv("C:/Users/EFergus/OneDrive - Environmental Protection Agency (EPA)/a_NLA_OE_project/Project_repository/Routput/Scenario_modeling/SEM_output/XERw_m3_EPT_CI.csv")
# Modify SEM output to get dotplot
m3_ept<- sem_ept%>%
  filter(str_detect(label,"_tot"))%>%
  rename(term=label,
         estimate=est.std,
         std.error=se,
         conf.low= ci.lower,
         conf.high=ci.upper,
         statistic=z,
         p.value=pvalue)%>%
  mutate(term = recode_factor(term, imp_tot="Impervious Rp",dam_tot="Dam",
                              precip_tot="Precipitation",phdi_tot="Drought index",
                              oebflow_tot="OE Bankfull flow",oelflow_tot="OE Summer flow",dexcess_tot="Evaporation indicator",
                              strpwr_tot="Specific strm power",
                              hag_tot="Agr index Rp", nhag_tot="Non-agr index Rp",rnat_tot="ForestGrass Rp",
                              xcmgw_tot="Riparian cover",rbs_tot="Bed stability", tn_tot="TN", sulf_tot="Sulfate"))%>%

  mutate(model="SEM")%>%
  select(term,estimate,std.error,statistic,p.value,model)

#m3_ept

#######################
# Bring EPT model output together
m_set_ept<-rbind(m1_ept, m2_ept,m3_ept)
coeff_dotplot_ept<-dwplot(m_set_ept)+
  theme_bw()+
  ggtitle("Std. model coeff: XER wade EPT")

coeff_dotplot_ept

# PRINT Coefficient Dot and whisker plot
# EPT XER
tiff(filename="C:/Users/EFergus/OneDrive - Environmental Protection Agency (EPA)/a_NLA_OE_project/Project_repository/Routput/Scenario_modeling/Figures/Dot_whisker_XER_EPT.tiff",
     width=5, height = 6, units="in", res=200)
coeff_dotplot_ept
dev.off()
```

Comparison of SEM and multiple regression approaches from the UCLA statistical consulting resource site that explains the differences we see in coefficient estimates and standard errors. 
https://stats.oarc.ucla.edu/r/seminars/rsem/#s2a

If I’m understanding it correctly, I think we are seeing differences for a few reasons. 
1)	The path analysis model allows for endogenous variables to predict other endogenous variables, whereas the multiple regression includes only exogenous variables. We could build a multivariate regression where only exogenous variables predict endogenous variables – and this would look like separate multiple regression models (if we fix covariances among endogenous variables to be zero – see point 3). But there’s good reason to include regression paths between endogenous variables based on our understanding of the system.
2)	The path analysis model we specify is not a saturated model meaning that we don’t have all possible arrows going to every box. Multiple regression models are saturated models meaning that all parameters are fully estimated and there are no degrees of freedom. Maybe if we made our path analysis model fully saturated, model coefficients would be similar to MLR. We could try this – but I think it would be easier to explore on a dummy, simplified model.
3)	The lavaan path analysis model by default includes covariances among the residual variances for any endogenous variables that do not have a regression path between one another (see diagram below). These covariances account for shared residual variance after accounting for specified predictor effects in the model. MLR assumes zero covariance. I think the covariance terms in the path analysis model lead to smaller standard errors around the model coefficients.
4)	In R lm() uses least squares and lavaan uses maximum likelihood for model estimation. The residual variance estimates will differ slightly using these different approaches.
